Studies of quantum dots

### 2016-03-25 Manuscript/paper.tex
Changed author list
spEnergiesQD/Manuscript/paper.tex: strip trailing whitespace
spEnergiesQD: Move extrapolation notes from paper.tex to extrapolation.txt

\author{Fei Yuan}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{Sarah Reimann}
\affiliation{Department of Chemistry and Center for Theoretical and Computational Chemistry, University of Oslo, N-0316 Oslo, Norway}
\author{Scott Bogner}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{Morten Hjorth-Jensen}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\affiliation{Department of Physics, University of Oslo, N-0316 Oslo, Norway}

\begin{abstract}
  We present and compare several many-body methods used to study a quantum dot
  system.  We calculate the approximate ground state energy of electrons in a
  circularly symmetric potential well using a harmonic oscillator basis
  optimized via Hartree-Fock.  We further improve the ground state energy
  using in-medium similarity renormalization group (IM-SRG).  Additionally, we
  calculate the addition and removal energies using quasidegenerate
  perturbation theory (QDPT).  Our results are benchmarked against quasi-exact
  diffusion Monte Carlo results from previous works.  Possible explanations
  for the discrepancies are discussed.
\end{abstract}

@@

The Hilbert spaces MR and PR is much closer in dimensionality than MR and PNP R . We
will therefore assume that we can use the same error formula for the spaces MR and PR .
As we will see later, this formula seems to hold very well for the systems we are studying.
We will also assume that ν(R) is negligible and rewrite Eq. (5.7) on the form
∞
∆E ≈ b ∑ (NP + r) r−c , b ∈ R.
(5.10)
r=R+1
This assumption is justified by Kvaal’s [16] results. He has performed FCI calculations on
two dimensional quantum dots in the energy cut basis with 3, 4 and 5 particles, and for these
systems ν(R) is shown to be small. Also, since PR ⊂ MR , we expect ν(R) to decay even
faster in the direct product basis.
5.3
Extrapolation Formulas
We have used Eq. (5.10) to parametrize the energy as a function of the number of shells.
R
E(R) ≈ a − b ∑ (NP + r)r− c ,
(5.11)
r=1
where a, b, c are real constants. This formula has an error ∼ E(R)ν(R), and consequently
it will become more accurate as R increases and ν(R) becomes small. This means that we
ˆ ˆˆ
should avoid extrapolations with energies ⟨PR H PR ⟩ for small R.
We also want to find the extrapolated energy at R → ∞ given that we know the optimal
values of a, b, c. We notice that E(∞) can be written in terms of the Riemann zeta function
ζ(c) such that
lim E(R) ≈ a − b[NP ζ(c) + ζ(c − 1)],
R→∞
where ζ(c) = ∑∞ r−c , which is known to converge for all c > 1.
r=1
34

In this section, we will calculate a set of extrapolated energies and compare these to
Diffusion Monte Carlo (DMC) energies which are not affected by a basis incompleteness
error. We use the extrapolation formula Eq. (5.11)
R
E(R) ≈ a − b ∑ (N + r)r−c ,
(25.1)
r=1
where N is the number of particles, R + 1 is the number of shells and a, b, c ∈ R are
constants. A curve fit of the FCIQMC values to E(R) yields the optimal parameters a, b, c.
The extrapolated energy is taken at the limit R → ∞, and the error is estimated using Eqs.
(13.18) and (13.22) assuming that Eq. (25.1) is exact.
As we discussed earlier, the extrapolation formula has an uncontrolled error ∝ ν(R)
where ν(R) is an unknown function. ν(R) is however assumed to fall off quickly as a
function of R. This is justified by the results of Kvaal [16] together with general consid-
erations about the form of the function. As we will see in this section, this assumption
seems to hold well for the systems that we have threated. To minimize ν(R) and reduce the
extrapolation error, we have only used energies calculated with R ≥ 5.

>>>>

\author{Fei Yuan}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{J\o rgen H\o gberget}
\affiliation{Department of Physics, University of Oslo, N-0316 Oslo, Norway}
\author{Sam Novario}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{Nathan Parzuchowski}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{Sarah Reimann}
\affiliation{Department of Chemistry and Center for Theoretical and Computational Chemistry, University of Oslo, N-0316 Oslo, Norway}
\author{Scott Bogner}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\author{Morten Hjorth-Jensen}
\affiliation{National Superconducting Cyclotron Laboratory and
  Department of Physics and Astronomy,
  Michigan State University,
  East Lansing, MI 48824, USA}
\affiliation{Department of Physics, University of Oslo, N-0316 Oslo, Norway}

\begin{abstract}
  We present and compare several many-body methods used to study a quantum dot
  system.  We calculate the approximate ground state energy of electrons in a
  circularly symmetric potential well using a harmonic oscillator basis
  optimized via Hartree-Fock.  We further improve the ground state energy
  using in-medium similarity renormalization group (IM-SRG).  Additionally, we
  calculate the addition and removal energies using quasidegenerate
  perturbation theory (QDPT).  Our results are benchmarked against quasi-exact
  diffusion Monte Carlo and Coupled Cluster calculations.
\end{abstract}

@@
